Evolutionary stretegies

### 2016-01-24 evolutionaryComputation/summery.tex
CHEATSHEET FINAL FORM

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-1explus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

@@

\begin{document}
\raggedright
\footnotesize
\begin{multicols}{3}

@@

\section{Genetic algorithms}
%TODO: crossover methods I don't understand
%TODO: schemas, buidling block theory
\subsection{Schema}
$o(h)$: schema order $o(11\#\#0)$ ie nonwildcard \# elements.
$\delta(h)$: schema defining length, maximum distance between two defining
symbols. $m(h,t)$: number of schema h instances at generation t.
$f(h,t)=\sum_{i\in P}f_i$ schema fitness is average fitness of members.

\emph{Buidling blocks} low order high performance schemate receive exponentiaonally
(geometrically) increasing trials.

@@

\section{Multi objective EA's}
\subsection{Pareto}
\emph{Dominance} A solution $x^0$ is said to (Paretor) \emph{dominate} a
solution $x^1$ (denoted $x^0\succ x^1$) iff
$(\forall i \in \{0,1,\dots,m-1\}:f_i(x^0)\leq f_i(x^1))\wedge
 (\exists i \in \{0,1,\dots,m-1\}:f_i(x^0)< f_i(x^1))$.
A solution is pareto optimal iff $\neg\exists x^1:x^1\succ x^0$.
The parote optimall set: $P_s=\{x^0|\neg\exists x^1:x^1\succ x^0\}$.
The paroto optimal front is the $P_f$ of all objective function values
correspoinding to the solutions in $P_s$: $P_F=\{f(x)=(f_0(x),f_1(x),\dots,
	f_{m-1}(x))|x\in P_s$
\subsection{Weighted aggregation}
The graphs with the linerider lines. It doesn't work with bowl
like objectives because only the edges are found.
\subsection{Evolutionary Multi-objective Optimization}
Desired result: representative subset of $P_s$.
\emph{Domination count}
For each solution count how many times
its dominatad. Then use one-dimensional selection (truncate or tournament).
\emph{Domination rank} solutions of rank 0 have domination count 0, solutions
of rank i are solutions of rank 0 when solutions of lower ranks are disregarded.

@@

\section{Model-based EAs: Est. of dist. algorithms}
Model based evolutionary algorthms: Population based stochastic search
algorithms. Exploitation: selection, Exploration: Learn model from selected
solutions, generate new solutions from the model (\&pop).
\subsection{Univariate PMBGA}
No statistical interaction between variables considered.
Model: probability vector $[p_1,\dots,P_l]$($l$:string length).
$p_i$: probability of value 1 at string position $i$.
$p(x)=\Pi^l_{i=1}p(x_i) \quad$ ($p(x_i)$: univariate marginal distribution).
Learn modal: count proportions of 1 in selected population. Sample model:
Generate new solutions with specified probabilities.

@@

\end{multicols}
\end{document}

>>>>

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-0.8ex plus -.5ex minus -.2ex}%
                                {0.2ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-0ex plus -.5ex minus -.2ex}%
                                {0.1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

@@

\begin{document}
\raggedright
\footnotesize
\begin{multicols}{4}

@@

\section{Genetic algorithms}
Darwinian system $\Rightarrow$ GA: 
Information structures $\Rightarrow$ binary strings, real-valued vectors,
programs etc. Structures are copied $\Rightarrow$ Selection algorithm.
Copies partially vary from original $\Rightarrow$ mutation \& crosover
operators. Structures are competing for a limited resource $\Rightarrow$
selecting fixed size parent pool. Reproductive succes depends on env
$\Rightarrow$ user defined fitness function.
\subsection{Schema}
$o(h)$: schema order $o(11\#\#0)$ ie nonwildcard \# elements.
$\delta(h)$: schema defining length, maximum distance between two defining
symbols. $m(h,t)$: number of schema h instances at generation t.
$f(h,t)=\sum_{i\in P}f_i$ schema fitness is average fitness of members.

\emph{Buidling blocks} low order high performance schemate receive exponentiaonally
(geometrically) increasing trials. ie there exists some schemas that when combined
will produce a near optimal solution, the building blocks.

@@

\section{Multi objective EA's}
\subsection{Pareto}
\emph{Dominance} A solution $x^0$ is said to (Paretor) \emph{dominate} a
solution $x^1$ (denoted $x^0\succ x^1$) iff
$(\forall i \in \{0,1,\dots,m-1\}:f_i(x^0)\leq f_i(x^1))\wedge
 (\exists i \in \{0,1,\dots,m-1\}:f_i(x^0)< f_i(x^1))$.
A solution is pareto optimal iff $\neg\exists x^1:x^1\succ x^0$.
The parote optimall set: $P_s=\{x^0|\neg\exists x^1:x^1\succ x^0\}$.
The paroto optimal front is the $P_f$ of all objective function values
correspoinding to the solutions in $P_s$: $P_F=\{f(x)=(f_0(x),f_1(x),\dots,
	f_{m-1}(x))|x\in P_s$
\subsection{Weighted aggregation}
The graphs with the linerider lines. It doesn't work with bowl
like objectives because only the edges are found.
\subsection{Multi-objective Optimization}
Desired result: representative subset of $P_s$.
\emph{Domination count}
For each solution count how many times
its dominatad. Then use one-dimensional selection (truncate or tournament).
\emph{Domination rank} solutions of rank 0 have domination count 0, solutions
of rank i are solutions of rank 0 when solutions of lower ranks are disregarded.

@@

\section{Model EAs: Est. dist. algorithms}
Model based evolutionary algorthms: Population based stochastic search
algorithms. Exploitation: selection, Exploration: Learn model from selected
solutions, generate new solutions from the model (\&pop).
\subsection{Univariate PMBGA}
No statistical interaction between variables considered.
Model: probability vector $[p_1,\dots,P_l]$($l$:string length).
$p_i$: probability of value 1 at string position $i$.
$p(x)=\Pi^l_{i=1}p(x_i) \quad$ ($p(x_i)$: univariate marginal distribution).
Learn modal: count proportions of 1 in selected population. Sample model:
Generate new solutions with specified probabilities.

@@

\subsubsection{COMIT}
Optimal dependency tree. Compute fully connected weighted graph between problem
variables. Weights are the mutual infromation $I(X,Y)$ between the variables.
Commit computes the maximam spanning treee of the weighted graph. The approximating
probability model is restricted to factorization in which the conditional
probabilty distribtuion for any random vraiable depends on the value of at
most one other random variable
$p(x)=\overset{n}{\underset{i=1}{\Pi}}p(X_i|X_{parent(i)})$.
$p(X)$ is the class of distubtions with a tree as graphical model.
The optimal tree model: Create a fully connected weighed graph $G$,
each vertix $V_i$ corresponds to randome variable $X_i$ each weight $W_{ij}$
for the edge betweeen $V_i$ and $V_j$ is equal to the mutal information
$I(X_i, X_j)$. The edges in the maximum spanning tree of $G$ determine an
optimal set $n-1$ conditional probabilites with which to aproximate the joint
probability distribution. Note that all ordered trees conforming the unorderd
spanning tree model identical distributions.
\emph{Algorithm} Calculate unconditional and conditinoal probabilties
$p(X_i)$ and $p(X_i,X_j)$, and the mutual information $I(X_i,X_j)$.
Select arbitrary random variable $X_r$ as root of the tree. Finda the
pair of variables $X_{in}$ and $X_{out}$ where $X_{in}$ is already in the
tree and $X_{out}$ is not, with the maximum mutual information
$I(X_{in}, X_{out})$. Add $X_{out}$ to the tree with $X_{in}$ as parent,
reapeat untill all variables are in the tree.

\subsection{Multivariate PMBGA}
MDL is a measure of complexity (Information Theroy).
$MDL(M,D)=D_{Model}+D_{Data}$ Model complexity $D_{Model}$:
complexity of describing the model. Compressed population complexity
$D_{Data}$: complexity of describing the data within the model (=measure
of goodness of the probability distribution estimation).
Best model = the one with the lowest MDL score.
$D_{Model}=log_2(N+1)\sum_i(2^{s_i}-1)$. Compressed pop complexity
$D_{Data}=N\sum_i H(M_i)$. $N$ pop size. $S_i$ size of partition $i$.
$M_i$ marginal distribution of the partition $i$. $H(M_i)$ entropy of the
marginal distribution of the partition i.

\subsubsection{Extended compact GA}
Greedily searches for the marginal product model that minzimizes the MDL.
$p(X)=\Pi^G_{g=1}p(x_g)$ Choose the probability distribution with the lowest
MDL score. Start from the simplest model: the univariate factorization.
Join two grousp that result in the largest improvemnt in the used scoring
measure. Stop when no jioning of two groups impvoes the score further.
\section{Model EAs: Linkage Tree GA}
Identify groups of problem variables that together make an important
contribution to the quality of solutions.
These variable groups are interacting in a non-linear way and should
be processed as a block = \emph{building block} the dependency struct
is a set of subsets of the problem.
\subsection{Family of subsets (dep struct)}
Let there be $l$ problem vars $x_0,\dots,x_{l-1}$. Let $S$ be a set of all
variable indecies $\{0,\dots,l-1\}$. A FOS $F$ is a set of subsets of the
set $S$. FOS $F$ i a subset of the powerset of $S(F\subseteq P(S))$.
FOS can be written more specifically as $F=\{F^0,\dots,F^{|F|-1}\}$. where
$F^i \subseteq \{0,\dots,l-1\}, i\in\{0,\dots,|F|-1\}$. Every var is
in at least one subset in the FOS: $\forall i \in \{0,\dots,l-1\}:(
\exists j \in \{0,\dots,|F|-1\}:i\in F^j)$.
\subsection{Linkage tree learning}
start from univariate structure (ie all vars in independent sets). Build
linkage tree using bottum-up hierarchial clustering algortim.
Similiartiy measuer: Between individual vairables X and Y: Mutual
informaiton I(X,Y).  Between clusture groups $X_{F^i}$ and $X_{F^j}$:
average pairwise linkage clustering (=unweighted pair group method
with arithmic mean).
$I^{UPGMA}(X_{F^i}, X_{F^j} = \frac{1}{|X_{F^i}||X_{F^j}|}
\sum\limits_{X\in X_{F^i}}\sum\limits_{Y\in X_{F^i}} I(X,Y)$
\subsubsection{Reciprocal nearest neighbor chain clustering}
Construct a nearsest neighbor chian: start from an arbitrary
start cluster which is followed by its nearest neigbro, which
again is followed by its nearest neigbor from among the remaining
points and so on.
When the nearest neigbor for the last chain element $n$ is the elements
$n-1$ we stop with the generation with the NNchan, we have found
a $RNN$ pair and can agglomerate it.
The nearest neigbor assignment stay valid for the remaining chain members,
which can thus be reused for the next iteration of agglomeration. If
we do not find any more RNN pairs in the remaining NN chain we create a new one
starting from a random start point.
\subsection{Optimal Mixing Evolutionary Algorhms}
OMEA is a model building EA that uses a FOS as its linkage model.
It uses intermediate function evalutations (inside variation).
Can be regarded as greedy improvment of exisiting solutions.
Coined ``optimal'' mixing because better instances for substructures
are imediatly accepted and not dependent on noise coming form other parts
of te solution.
\subsection{GOMEA}
FOS liknage models to specify the linked vars. A subset of the fos is used
as crosover mask. Crosover is greedy: only improvments (or equal) are accepted.
Each generation a new FOS model is build from selected solutions. For each solution
in the population all subsets of the FOS are treid with a donor solution
randomly picked fro mthe population. Recombainative OM is GA-like: select single
donor solution to perform OM with. GOM is EDA-like: select donor solution for each
subset in OM.
\subsection{Linkage Tree genetic algorithm.}
The LTGA is an instance of GOMEA that uses a linkage tree as fos model. Each
generation a new hierechial cluster tree is build. For each solution in
population, traverse tree starting at the top. Node (=clusters) in the linkage
tree used as crosover masks. Select random donor solution and its values at
the crossover mask replace the variable values from the current solution.
Evaluate new solution and accept if better/equal, otherwise reject.
\section{Seminars}
\subsection{ILS for Qaudratic assignment problem}
Population based extension work by replacing the worst. Using evolution
strategies. and don't do recombination?
\subsection{GA with ILS for solving location routing}
Assignment Configuration $A(x)$: location. Permutation vector $P(x)$: routing.
\emph{crossover} $A(x)$ one point cross. $P(x)$ focus on relative ordering.
\subsection{GASAT}
\emph{Corrective clause crossover} set all the bits of the child to undefined.
Go over all false clauses of both parents (and child), find the best to flip,
keep the sum of parents' fitness the lowest, flip this bit in the child.
End step: all undefined bits of the child are uniformly crossed over.
\emph{Corrective clause and truth maintance crossover} Obvious problem
with CC: Child may make clause false that is true for both parents. A possible
olution: CCTM. DO CC except for end step. Go over all clauses that are true
for parents, false for child. Look at the bits of the clause but only those
different from the child and at least one parent. Find the best bit to flip,
keep the sum of the parents' fitness the lowest. Flip this bit in the cild.
End step of CC.
\emph{Fleuent and ferland's crossover} Go over all clauses true for one parent
and false for another. Copy all of the former's bits for that clause. End
step of CC.
\emph{Also uniform crossover} copy a bit of x or y with equal probablity.
\subsection{Fitness landscapes in graphbitpart}
Researh suggest that the optimal soution is often found near the
center of local-optimum space. Thus a good idea for a search
tactic is to find a way to get near this center and use a lcoal
optimization function to try and get near the global optimum.
\subsection{Solving bus terminal problems}
Standar rollete wheel with linear ranking is assigning fitness to
possible solutions. The higher the fitness the more chance of being
selected.
\subsection{GLS w. dist preserv recombo}
Order crossover (see earlier). single parent cross, select with probability 0.5
from the parent for customers not inserted pick one at random and insert at the
best position. Convert into chromosone.
String cross
\end{multicols}
\end{document}
